---
title: "Get ssb data"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup}
library(PxWebApiData)
library(tidyverse)
library(readxl)
library(sf)
library(sp)
library(spdep)
library(spatialreg)
library(broom)
library(mctest)
library(huxtable)
library(tmap)
```

# Read in map 2017 (unedited)

The map is converted from 432 to 426 multipolygons in Qgis (Vektor > Geometriverktøy > Samle geometrier ...). The routine with " group_by(knr) %>%  summarise(geometry = st_combine(geometry))" did not work.

```{r}
map_2017 <- st_read("map_2017_426.gpkg")
map_2017 <- map_2017 %>% 
  rename(
    knr = KOMMUNENUM
  )

```

# Get knr and kNavn from local file

We want the Norvegian names, hence we have to adjust the municipality names for Vetsfold 2017.

```{r}
load("bld06_19.Rdata")

knr_and_knavn <- bld90_19 %>% 
  select(knrNavn) %>% 
  unique() %>%
  # separate at the space after last digit
  separate(col = knrNavn, into = c("knr", "kNavn"), sep="\\s", extra = "merge") %>% 
  filter(!knr %in% c("0712", "0715", "0729")) %>% 
  bind_rows(
    c(knr = "0702", kNavn = "Holmestrand"),
    c(knr = "0709", kNavn = "Larvik"),
    c(knr = "0714", kNavn = "Hof"),
    c(knr = "0722", kNavn = "Nøtterøy"),
    c(knr = "0723", kNavn = "Tjøme"),
    c(knr = "0728", kNavn = "Lardal")
  )
```


# Get ssb data via api

This document is all about getting data registered on the municipality level from SSB (one trick to get a list of all these municipality tables is to search at SSB for `(K)` and then click on `Statistikkbanken` in the left column). We will use the SSB api exposed through the `PxWebApiData` R package. This API can also be used against most Nordic SSB sister organisations. Since we have a list of the relevant municipalities for 2017 (`map_2017$knr`) we can restrict our api request to just these. We are just going to use some of these variables in the following example, but the other variables serves as examples and some might be used in further analysis. There is still plenty of relevant data at SSB that it should be quite trivial to get access to using the techniques shown below.

New dataset. Use pm2 for 2017 to avoid the big changes in Trøndelag 1.1 2018.

## Get avarage price per sqr. meter single family dwellings in each municipality (426)

```{r}
pm2_2017 <- ApiData(
    urlToData = "06035",
    Region = map_2017$knr,
    Boligtype = "01",
    Tid = "2017"
)

pm2_2017 <- pm2_2017$dataset

# list with 2 elements, we want dataset
pm2_2017 <- pm2_2017 %>% 
  select(-Boligtype, -Tid) %>% 
  rename(
    knr = Region
  ) %>% 
  pivot_wider(id_cols = knr, names_from = "ContentsCode", values_from = "value") %>% 
  rename(
    pm2 = KvPris,
    num_dwelling_sales = Omsetninger 
  )

# We are tidy
dim(pm2_2017)[1]
```

```{r}
# 139 missing values. Try other years for these
pm2_missing <- pm2_2017 %>% 
  filter(is.na(pm2)) %>% 
  select(knr)

pm2_15to18 <- ApiData(
    urlToData = "06035",
    Region = pm2_missing$knr,
    Boligtype = "01",
    Tid = c(as.character(2010:2016), "2018")
)

tmp <- pm2_15to18$dataset %>% 
  select(-Boligtype) %>% 
  rename(knr = Region) %>% 
  pivot_wider(id_cols = knr, names_from = c("ContentsCode", "Tid"), values_from = "value") %>% 
  mutate(
    pm2 = rowMeans(select(., KvPris_2010:KvPris_2018), na.rm = TRUE),
    num_dwelling_sales = rowMeans(select(., Omsetninger_2010:Omsetninger_2018), na.rm = TRUE)
  ) %>% 
  select(knr, pm2, num_dwelling_sales)

# Verty crude, we just take the mean of existing values

pm2_2017 <- pm2_2017 %>% 
  left_join(tmp, by = "knr") %>% 
  mutate(
    pm2 = if_else(is.na(pm2.x), as.integer(pm2.y), pm2.x),
    num_dwelling_sales =  if_else(is.na(pm2.x), as.integer(num_dwelling_sales.y), num_dwelling_sales.x)
    ) %>% 
  select(knr, pm2, num_dwelling_sales)

rm(tmp)
```




## Population, three categories 0-19, 20-64 and 65+

```{r}
# Kjonn = FALSE, aggregate F+M
# The F in "F0-19" DOES NOT indicate Female
bef_2017 <- ApiData(
    urlToData = "07459",
    Region = map_2017$knr,
    Alder = list("agg:TredeltGrupperingB2", 
                 c("F0-19", "F20-64", "F65+")),
    Kjonn = FALSE,
    Tid = "2017"
)

# list with 2 elements, we want dataset
bef_2017 <- bef_2017$dataset
```

```{r}
head(bef_2017, n=6)

bef_2017 <- bef_2017 %>% 
  select(Region, Alder, value) %>% 
  rename(
    alder = Alder,
    knr = Region,
    bef = value,
  )

bef_2017 <- bef_2017 %>% 
  pivot_wider(id_cols = knr, names_from = "alder", values_from = "bef") %>% 
  # rename so we do not have to escape names
  rename(
    pop_0_19 = `F0-19`,
    pop_20_64 = `F20-64`,
    pop_65p = `F65+`
    ) %>% 
  # Calculate tot. pop.
  mutate(tot_pop = pop_0_19 + pop_20_64 + pop_65p)

# we are tidy
head(bef_2017, n = 5)
dim(bef_2017)[1]
```

## Education level (percentage each level)

```{r}
# Get metadata 
#ApiData("http://data.ssb.no/api/v0/no/table/09429",  returnMetaFrames = TRUE)
```

```{r}
edu_2017 <- ApiData(
    urlToData = "09429",
    # we want in percent
    ContentsCode = "PersonerProsent",
    Region = map_2017$knr,
    # Drop "00" since total, 100%, also dro "09a" Uoppgit
    Nivaa = c("01", "02a", "11", "03a", "04a"),
    Kjonn = FALSE,
    Tid = "2017"
)

edu_2017 <- tibble(
  # want knr
  knr = edu_2017$dataset$Region , 
  # want level descriptions, not code
  nivaa = edu_2017$`09429: Personer 16 år og over, etter region, nivå, statistikkvariabel og år`$nivå,
  edu_lev_percent= edu_2017$dataset$value
  ) 

head(edu_2017, n=7)

#make tidy
edu_2017 <- edu_2017 %>% 
   pivot_wider(
     id_cols = knr, 
     names_from = "nivaa", 
     values_from = "edu_lev_percent"
     ) %>% 
  rename(
    edu_prim = `Grunnskolenivå`,
    edu_sec = `Videregående skolenivå`,
    edu_voc = `Fagskolenivå`,
    edu_uni_short = `Universitets- og høgskolenivå, kort`,
    edu_uni_long = `Universitets- og høgskolenivå, lang`
  )

# Tidy OK
head(edu_2017)
dim(edu_2017)[1]
```

## Household income

```{r}
# Get metadata 
#ApiData("http://data.ssb.no/api/v0/no/table/07183",  returnMetaFrames = TRUE)
```

```{r}
inc_2017 <- ApiData(
    urlToData = "07183",
    # "Hushold" would also get us total number of households in muni
    ContentsCode = c(paste("Inntekt", seq(from = 1, to = 7), sep = "")),
    Region = map_2017$knr,
    Tid = "2017"
)

head(inc_2017$`07183: Husholdninger, etter region, statistikkvariabel og år`, n=9)
```


```{r}
inc_2017 <- tibble(
  # want knr
  knr = inc_2017$dataset$Region , 
  # want level descriptions, not code
  inc_cat = inc_2017$`07183: Husholdninger, etter region, statistikkvariabel og år`$statistikkvariabel,
  inc_lev_percent= inc_2017$dataset$value
  )

inc_2017 <- inc_2017 %>% 
  pivot_wider(
     id_cols = knr, 
     names_from = "inc_cat", 
     values_from = "inc_lev_percent"
     ) %>% 
  rename(
    # Some problem with names here, had to do names(inc_2017) above and 
    # then copy the text strings
    inc_0_150 = `Samlet inntekt under 150 000 kr, prosent`,
    inc_150_250 = `Samlet inntekt 150 000 - 249 999 kr, prosent`,
    inc_250_350 = `Samlet inntekt  250 000 - 349 999 kr, prosent`,
    inc_350_450 = `Samlet inntekt  350 000 - 449 999 kr, prosent`,
    inc_450_550 = `Samlet inntekt  450 000 - 549 999 kr, prosent`,
    inc_550_750 = `Samlet inntekt  550 000 - 749 999 kr, prosent`,
    inc_750k_p = `Samlet inntekt 750 000 kr og over, prosent`
  )
# we are tidy
head(inc_2017, n=3)
dim(inc_2017)[1]
```

## Household net capital (wealth)

```{r}
wealth_2017_ssb <- ApiData(
    urlToData = "10320",
    # "Hushold" would also get us total number of households in muni
    ContentsCode = c("Hushald (prosent)"),
    Nettoform = c("12", "13", "10", "25", "31", "11", "32"),
    Region = map_2017$knr,
    Tid = "2017"
)
#head(wealth_2017$dataset, n=10)
#head(wealth_2017$`10320: Hushald (prosent), etter region, nettoformue, statistikkvariabel og år`, n=10)
```

```{r}
wealth_2017 <- wealth_2017_ssb$dataset %>% 
  tibble() %>% 
  select(-ContentsCode, -Tid) %>% 
  rename(
    knr = Region,
    net_cap = Nettoform,
    cap_lev_percent = value) %>% 
  pivot_wider(
     id_cols = knr, 
     names_from = "net_cap", 
     values_from = "cap_lev_percent"
     ) %>% 
  rename(
    net_cap_0_250k = `12`,
    net_cap_250k_500k = `13`,
    net_cap_500k_1000k = `10`,
    net_cap_1000k_2000k = `25`,
    net_cap_2000k_3000k = `31`,
    net_cap_3000k_4000k = `11`,
    net_cap_4000k_p = `32`,
  )

# we are tidy
head(wealth_2017, n=3)
dim(wealth_2017)[1]
```
## Employed by government

```{r}
gov_emp_2017 <- ApiData(
    urlToData = "12628",
    # total num.of gov. emp. in each muni
    StatligEnhet = "0",
    # in yearly eq.
    ContentsCode = "AvtalteArsverk",
    Region = map_2017$knr,
    Tid = "2017"
)

gov_emp_2017 <- gov_emp_2017$dataset %>% 
  select(Region, value) %>% 
  rename(
    knr = Region,
    gov_emp = value
  )
#we are tidy
head(gov_emp_2017, n=5)
dim(gov_emp_2017)[1]
```

## Wholesale and retail trade sales

```{r}
sales_per_capita_2017  <- ApiData(
    urlToData = "04776",
    Region = map_2017$knr,
    Tid = "2017"
)

sales_per_capita_2017  <- sales_per_capita_2017$dataset %>%  
  select(Region, value) %>% 
  # spc sales per capita
  rename(
    knr = Region,
    spc = value
  )
# we are tidy
head(sales_per_capita_2017)
dim(sales_per_capita_2017)[1]
```

## Household income before and after tax in decile (percent)

Percent relative to national decile.

```{r}
inc_dec_2017 <- ApiData(
    urlToData = "12558",
    ContentsCode = "AndelHush",
    Desiler = c(paste(0, 1:9, sep=""), "10"),
    Region = map_2017$knr,
    Tid = "2017"
)

inc_dec_2017 <- tibble(
  # want knr
  knr = inc_dec_2017$dataset$Region , 
  # want level descriptions, not code
  tax = inc_dec_2017$`12558: Husholdninger, etter region, inntekt før/etter skatt, desil, statistikkvariabel og år`$`inntekt før/etter skatt`,
  decile = inc_dec_2017$dataset$Desiler,
  percent = inc_dec_2017$dataset$value
  )

inc_dec_2017 <- inc_dec_2017 %>% 
  mutate(
    tax = case_when(
      # income before and after tax
      tax == "Samlet inntekt" ~ "inc_bt",
      tax == "Inntekt etter skatt" ~ "inc_at",
      TRUE ~ as.character(tax)
    )
  ) %>%
  pivot_wider(
     id_cols = knr, 
     names_from = c("tax", "decile"), 
     values_from = "percent"
     )

# we are tidy
head(inc_dec_2017, n=2)
dim(inc_dec_2017)[1]
```

## Industries

```{r}
num_emp_2017_ssb <- ApiData(
    urlToData = "07196",
    OrgFormer = c("99","02","03","04","10","24"),
    AntAnsatte = c("99", "00", "01","02","03","04","05","15","16"),
    NACE2007 = c("01-99"),
    Region = map_2017$knr,
    Tid = "2017"
)
#so that we keep the ssb data
num_emp_2017 <- num_emp_2017_ssb$`07196: Foretak (1.1.) unntatt offentlig forvaltning og primærnæringene, etter region, organisasjonsform, næring (SN2007), antall ansatte, statistikkvariabel og år`
num_emp_2017$knr <- num_emp_2017_ssb$dataset$Region

num_emp_2017 <- num_emp_2017 %>% 
  # drop two variables
  select(-region, -statistikkvariabel, -år, -`næring (SN2007)`) %>% 
  # want knr to the left, really not necessary 
  select(knr,everything()) %>% 
  # rename before pivot to names liked by R
  rename(
    org_form = organisasjonsform,
    num_emp_group = `antall ansatte`,
    firms = value
  )

num_emp_2017 <-  num_emp_2017 %>% 
  pivot_wider(
     id_cols = knr, 
     names_from = c("org_form", "num_emp_group"), 
     values_from = "firms"
     )
# we are tidy
head(num_emp_2017)
dim(num_emp_2017)[1]
```

## Supply of single family dwellings

```{r}
sup_sfd_2017 <- ApiData(
    urlToData = "05940",
    Byggeareal = c("111", "112"),
    Region = map_2017$knr,
    Tid = "2017"
)

sup_sfd_2017 <- sup_sfd_2017$`05940: Boligbygg, etter region, bygningstype, statistikkvariabel og år` %>% 
  as_tibble() %>% 
  mutate(region = sup_sfd_2017$dataset$Region) %>% 
  rename(knr = region) %>% 
  select(-år) %>% 
  unite(tmp, bygningstype, statistikkvariabel, sep="_") %>% 
    pivot_wider(
     id_cols = knr, 
     names_from = "tmp", 
     values_from = "value"
     ) %>% 
  # sfd Single Family Dwelling
  # We have to escape the old variable names because of the spaces 
  rename(
    sfd_started = `Enebolig_Igangsatte boliger`,
    sfd_ba_started = `Enebolig_Bruksareal igangsatte boliger (m²)`,
    sfd_ba_completed = `Enebolig_Bruksareal fullførte boliger (m²)`,
    # w2u with 2 units
    sfd_w2u_started = `Enebolig med hybelleilighet, sokkelleilighet o.l._Igangsatte boliger`,
    sfd_w2u_ba_started = `Enebolig med hybelleilighet, sokkelleilighet o.l._Bruksareal igangsatte boliger (m²)`,
    sfd_w2u_ba_completed = `Enebolig med hybelleilighet, sokkelleilighet o.l._Bruksareal fullførte boliger (m²)`  
    )

# we are tidy
head(sup_sfd_2017, n=4)
dim(sup_sfd_2017)[1]
```



## Centrality index

From SSB, for the year 2018, but we use it for 2017. Probably very small changes from 2017 to 2018. Changed knr for Trøndelag to the old numbers. Split 5054 Indre Fosen into 1624 Rissa and 1718 Leksvik. The last two was given the same centrality as the combined entity.


```{r}
centrality_2018_ed <- read_excel("Sentralitet 2018_ed.xlsx")
#View(centrality_2018_ed)
```



## Put it all together

```{r}
names(pm2_2017)

data_2017 <- knr_and_knavn %>% 
  left_join(pm2_2017) %>% 
  left_join(bef_2017) %>% 
  left_join(edu_2017) %>% 
  left_join(inc_2017) %>% 
  left_join(gov_emp_2017) %>%
  left_join(sales_per_capita_2017) %>%
  left_join(inc_dec_2017) %>%
  left_join(wealth_2017) %>% 
  left_join(num_emp_2017) %>%
  left_join(sup_sfd_2017) %>% 
  left_join(centrality_2018_ed ) %>% 
  mutate(
    log_pm2 = log(pm2),
    spc_100k = spc/100000,
    gov_emp_1k = gov_emp/1000,
    sen_ind_1k = sen_ind/1000
  ) %>% 
  # rearrange 
  select(knr, kNavn, pm2, log_pm2, everything())

#View(data_2017[, c("knr", "sen_ind")])
```

```{r}
#head(data_2017, n=5)
dim(data_2017)
```

We have 115 variables and 426 observations.


## Add data to our map

```{r}
# join map and data
map_data_2017 <- map_2017 %>% 
  left_join(data_2017)
# write out new map with data
# if one add new variables it might be necessary to delete the file before writing new version
st_write(obj = map_data_2017, dsn = "map_data_2017.gpkg", driver = "GPKG", append=FALSE)
```

```{r}
#siste
```

